Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-02-23 ms.tex
magic

\subsection{Nuclear X-ray Luminosities of Quiescent Galaxies}
\label{sec:mdot}

Figure \ref{fig:vwSources} shows that the CNM heating rate is on
average approximately constant with SMBH mass, except for the largest
black holes with $M_{\bullet} \gtrsim 5\times 10^{8}M_{\odot}$.  For
fixed wind heating $\dot{M}/\dot{M}_{\rm edd}$ increases with SMBH
mass $\propto M_{\bullet}^{0.5(0.8)}$ for core(cusp) galaxies,
respectively (Fig.~\ref{fig:mdot_mass}; eq.~[\ref{eq:eddr_analytic}]).
This appears in contradiction with the nuclear X-ray luminosities of
quiescent galaxies, which (though showing broad scatter) instead on
average obey $\langle L_X/L_{\rm edd}\rangle \sim \Mbh^\alpha$, where
$\alpha\simeq -0.25$ \citep{Pellegrini10, Miller+15}.
Figure~\ref{fig:miller} shows our predicted value of $\langle
L_{X}/L_{\rm edd} \rangle =\epsilon_X \Mdot c^2$ as a function of
black hole mass for core (blue) and cusp (green) galaxies.  The left
panel assumes a constant low radiative efficiency $\epsilon_X =
10^{-4}$ typical of those estimated for low luminosity AGN (e.g.,
\citealt{DiMatteo+00}, \citealt{Ho:2009a}) and SgrA* (e.g.,
\citealt{Quataert+99}, \citealt{Quataert:2004a}), while the right
panel shows an efficiency of $\epsilon_X= \dot{M}/\dot{M}_{\rm edd}$
that depends on the Eddington ratio in the way predicted by simple
models for radiatively inefficient accretion flows
(\citealt{Narayan&Yi95}).  Superimposed with the theoretical curves
are symbols showing X-ray measurements (red stars) and upper limits
(red triangles) from \citet{Miller+15}, along with the best power-law
fit (red line; {\bf BDM: change to this!}) and the approximate X-ray
detection threshold (gray line).  Finally, shown with purple and
orange lines are the maximum accretion rates for thermally stable
accretion (eq.~[\ref{eq:Mdotmax}]) and set by Ia blow-out
(eq.~[\ref{eq:eddr_Ia}]), as well as the maximum accretion rate
allowed by Compton heating (eq.~[\ref{eq:MdotC}]).

For either of our assumptions about radiative efficiency our predicted
trend in $\langle L_{X}/L_{\rm edd}\rangle$ for core and cusp galaxies
is opposite the observationally inferred one.  However, the observed
$\langle L_{X}/L_{\rm edd}\rangle$ do imply the accretion rates of
observed galaxies are below the maximum ones for thermally stable
accretion if $\epsilon_X \lesssim 10^{-4}$.  This suggests that most
nearby galaxies are consistent with accreting in a thermally stable
steady state, even if our estimate of their predicted heating does not
match the observed trend with $M_{\bullet}$ perfectly.  On the other
hand if $\epsilon_{X} \sim 10^{-5}$ then the trend would be consistent
with the accretion rate set by Ia blow-out.  Also note that in the
case of $\dot{M}-$dependent radiative efficiency ($\epsilon_X =
\dot{M}/\dot{M}_{\rm edd}$), the predicted branches of $\langle
L_{X}/L_{\rm edd}\rangle$ decreases with decreasing $M_{\bullet}$,
calling into question attempts to extrapolate the $\langle L_{X}
\rangle - M_{\bullet}$ relationship to low $M_{\bullet}$ in order to
constrain the occupation fraction of black holes in the lowest mass
galaxies (\citealt{Miller+15}).

>>>>

\subsection{Nuclear X-ray Luminosities}
\label{sec:mdot}

Figure \ref{fig:vwSources} shows that the average CNM heating rate is predicted to be approximately constant with SMBH mass, except for the largest black holes with $M_{\bullet} \gtrsim 5\times 10^{8}M_{\odot}$.  For
a fixed wind heating parameter, $\dot{M}/\dot{M}_{\rm edd}$ increases with SMBH
mass $\propto M_{\bullet}^{0.5(0.8)}$ for core(cusp) galaxies,
respectively (Fig.~\ref{fig:mdot_mass}; eq.~[\ref{eq:eddr_analytic}]).
This dependence appears in contradiction with the nuclear X-ray luminosities of
quiescent galaxies, which (though showing several orders of magnitude scatter) instead on average obey $\langle L_X/L_{\rm edd}\rangle \sim \Mbh^\alpha$, where $\alpha \in [-0.4,-0.2]$ \citep{Zhang+09, Pellegrini10, Miller+15}.

Figure~\ref{fig:miller} shows our predicted value of $\langle
L_{X}/L_{\rm edd} \rangle =\epsilon_X \Mdot c^2$ as a function of
black hole mass for core (blue) and cusp (green) galaxies.  The left
panel assumes a constant low radiative efficiency $\epsilon_X =
10^{-4}$ typical of those estimated for low luminosity AGN (e.g.,
\citealt{DiMatteo+00}, \citealt{Ho:2009a}) and SgrA* (e.g.,
\citealt{Quataert+99}, \citealt{Quataert:2004a}), while the right
panel shows an efficiency of $\epsilon_X= \dot{M}/\dot{M}_{\rm edd}$
that depends on the Eddington ratio in the way predicted by simple
models for radiatively inefficient accretion flows
(\citealt{Narayan&Yi95}).  Superimposed with the theoretical curves
are symbols showing X-ray measurements (red stars) and upper limits
(red triangles) from \citet{Miller+15}, along with the best power-law
fit (red line; {\bf BDM: change to this!}) and the approximate X-ray
detection threshold (gray line).  Finally, shown with purple, orange, and {\bf brown} lines
orange lines are the maximum accretion rates for thermally stable
accretion (eq.~[\ref{eq:Mdotmax}]), set by SN Ia blow-out
(eq.~[\ref{eq:eddr_Ia}]), and allowed by Compton heating feedback (eq.~[\ref{eq:MdotC}]), respectively.

First note that to first order, our predicted values of $\langle L_{X}/L_{\rm edd}\rangle$ for heating corresponding to average star formation reside reasonably close to the data $\langle L_{X}/L_{\rm edd}\rangle$.  However, for either assumption about radiative efficiency, our predicted positive trend of $\langle L_{X}/L_{\rm edd}\rangle$ with black hole mass is opposite the observationally inferred one, which instead shows $L_{X}/L_{\rm edd}$ decreasing with $M_{\bullet}$.   Comparison to the data is improved, however, by taking into account the fact that larger(smaller) black hole masses are more commonly associated with core(cusp) galaxies, resulting in a flatter predicted trend for $\langle L_X/L_{\rm edd}\rangle$ when both samples are combined.   The large scatter observed in $L_{X}/L_{\rm edd}$ of typically two to four orders of magnitude in the \citet{Pellegrini10} sample can also be understood in our steady-accreting model by the .  Alternatively, cases where the older stellar population precludes a steady accretion flow, the scatter could be due to (e.g.~\citealt{Pellegrini&Ciotti98}; refs).

   Finally, we note that \citet{Pellegrini10} finds that $L_{X}$ does not show any dependence on the age of the stellar population in the central galactic region for old populations with stellar ages $> 3$ Gyr, although less X-ray luminous nuclei are found among galaxies with a younger stellar population.  This

 however, when the

.  However, this conclusion comes with several caveats: (1)

However, the observed values of 
$\langle L_{X}/L_{\rm edd}\rangle$ do imply the accretion rates of
observed galaxies are below the maximum ones for thermally stable
accretion if $\epsilon_X \lesssim 10^{-4}$.  This suggests that most
nearby galaxies are consistent with accreting in a thermally stable
steady state, even if our estimate of their predicted heating does not
match the observed trend with $M_{\bullet}$ perfectly.  On the other
hand if $\epsilon_{X} \sim 10^{-5}$ then the trend would be consistent
with the accretion rate set by Ia blow-out.  Also note that in the
case of $\dot{M}-$dependent radiative efficiency ($\epsilon_X =
\dot{M}/\dot{M}_{\rm edd}$), the predicted branches of $\langle
L_{X}/L_{\rm edd}\rangle$ decreases with decreasing $M_{\bullet}$,
calling into question attempts to extrapolate the $\langle L_{X}
\rangle - M_{\bullet}$ relationship to low $M_{\bullet}$ in order to
constrain the occupation fraction of black holes in the lowest mass
galaxies (\citealt{Miller+15}).

Past studies have found that core galaxies show a wide range of nuclear radio and X-ray luminosities, while cusp galaxies are confined below a threshold (e.g., \citealt{Bender+89, Pellegrini99, Capetti&Balmaverde05}).
